Comment: Reasoning about Interference Between Units

### 2015-12-19 response.md

We really appreciate the constructive and timely feedback from you and the reviewers.

1) As Reviewer 1 suggests, how does the model perform as the assumptions
about spill overs deviate?  How sensitive are the inferences to the
particular assumptions you can make?

In the paper we clarified that we are focusing on testing sharp null
hypotheses arising from one toy model of how an experimental treatment might
propagate across a social network. We have added some discussion to the paper
highlighting the fact that the power of a test of a sharp null hypothesis
depends on *both* the model of causal effects *and* the test statistic (and
the effective sample size which itself depends on sample size and network
characteristics). We also note that, given randomization, any test statistic
ought to produce valid statistical inferences (i.e. tests that would reject
the truth no more than $100 \alpha \%$ of the time).

Since the point of the paper is to learn about test statistics,
we do not explore alternative models of causal effects. The original
paper (BFP) on this method compared the performance of tests when the data did
and did not arise from the causal model and we point to that paper for those
who would like to pursue those set of questions. We did add some discussion
about the power of the SSR test statistic against different kinds of causal
effects: the SSR test statistic ought to be particularly sensitive to changes
in mean differences (after all, recall that the coefficient on a dichotomous
variable in an least squares model with a continuous outcome that minimizes
the SSR is the difference of means) and ought to be less sensitive to other
differences.

@@

2) Reviewer 2 asks you to carry forward your toy example with four units
through the entire analysis.  If this isn't possible, please provide the
simplest case that illuminate the core intuition of your method.

When we set about adding the four unit example to the different sections of the
paper, we realized that we were expanding the length of the research note.

@@

3) Reviewer 2 also asks you to provide intuition about when the KS-test
would outperform the Sum of Squared Residuals.  As part of a simulation or
analysis of a simple case, demonstrating when/why the KS test outperforms
will be useful for gaining intuition about why the SSR is a generally
superior method.

In the text we note that the SSR is particularly sensitive to differences in
means (or perhaps symmetric differences in tails that also produce a lot of
movement in the mean of one group relative to another group) while the KS-test
ought to have more power to detect asymmetric differences in the distributions
of the treated versus control groups --- for example, differences that change
variance without moving the means by stretching one tail of the distribution.

4) In letters we have an emphasis that new methods matter when applied to
real examples.  Can you demonstrate that a case where the SSR will change
the inferences we make when applied to a real world example?

We have added a section in which we replicate the analysis in Coppock's 2014
*Journal of Experimental Political Science* article "Information Spillovers:
Another Look at Experimental Estimates of Legislator Responsiveness".

### 2015-12-19 paper/introduction.tex
More progress on editing

Notice that the test statistic choice matters in this process: the engine of
statistical inference involves summarizing information against the
hypothesized claim, yet different ways to summarize information might be more
or less sensitive to substantively meaningful differences. The statistical
power of a simple test of the sharp null hypothesis of no effects will vary as
a function of the design of the study (proportion treated, blocking structure,
etc), characteristics of the outcome (continuous, binary, skewed, extreme
points, etc), and the way that a test statistic summarizes the outcome (does
it compare means, standard deviations, medians, qqplots, etc). In general,
test statistics should be powerful against relevant alternatives. \citet[\S 2.4.4]{rosenbaum:2002} provides more specific
advice about the large sample performance of certain classes of test statistics
and BFP repeat his general advice: ``Select a test statistic [$\mathcal{T}$] that will be
small when the treated and control distributions in the adjusted data \ldots
are similar, and large when the distributions diverge.'' \cite[Proposition 4
and 5, \S 2.9]{rosenbaum:2002} presents results proving that test statistics
with this property (``effect increasing'' test statistics), produce an
unbiased test of the hypothesis of no effects or positive effects when the
positive effects involve one parameter. Notice that this means that, a test of
the sharp null of no effects
based on a known randomization using \emph{any} effect increasing
test-statistic to be valid test (in that the test should produce $p$-values
less than $\alpha$ no more than 100$\alpha$ of the time when the null is true) even though different
test statistics may imply different power against false hypotheses.  Yet, when the models are complex and
may involve increasing effects in the direction of one parameter and
non-linear effects in the direction of another parameter, BFP showed that
sometimes a KS-test will have (1) no power to address the model at all such
that all hypothesized parameters would receive the same implausibility
assessment; (2) or might reject all such hypothesized parameters. Thus,
although in theory one may assess sharp multiparameter hypotheses,
 in practice one may not learn much from such tests. BFP thus recommended simulation studies of the operating
characteristics of tests as a piece of their workflow --- because  the theory
justifying simple one-dimensional effect increasing test statistics clearly
did not cover multi-parameter situations like those easily arising from social
network experiments.

@@

Yet, one can also think about the process of hypothesis testing as a process
of assessing model fit, and there are usually better ways to evaluate the fit
of a model than comparing two marginal distributions. In the case where we know
the fixed adjacency matrix of the network, $\bS$, and where we imagine that
network attributes (like degree) of a node play a role in the mechanism by
which treatment propagates, the idea of assessing model fit rather than
closeness of distributions leads naturally to the the sum-of-squared-residuals
(SSR) from a least squares regression of $ \widetilde y_{i,Z_i=0}$ on $Z_{i}$ and
$\bz^{T} \bS$ (i.e.\ the number of directly connected nodes assigned treatment)
as well as the $\mathbf{1}^{T} \bS$ (i.e.\ the degree of the node). If we
collected $Z_{i}$,$\bz^{T} \bS$, and $\mathbf{1}^{T} \bS$ into a matrix $\bX$,
and fit the $\widetilde y_{i,Z_i=0}$ as a linear function of $\bX$ with
coefficients $\bbeta$ when we could define the test statistic as:

\begin{equation}
 \T(\yu,\bz)_{\text{SSR}} \equiv \sum_{i} ( \widetilde y_{i,Z_i=0} -
 \bX\hat{\bbeta}) ^2 \label{eq:ssr}
\end{equation}

@@

\citet{coppock2014information} presents an reanalysis of an experiment performed
by \citet{butlernickerson2011}. Leading up to a key vote on a budget bill, SB24, in the
New Mexico state senate, legislators were randomly assigned to see
constituent survey results, and their votes on SB24 recorded. The original
analysis found that legislators assigned to receive constituent information were
more likely to vote consistently with the preferences of their district.
Given the easy nature of sharing information in
field experiments \citep{Winters:2012fk}, Coppock reanalyzes the
experimental results with an eye towards spillover, constructing a social
network based on the similarity of W-NOMINATE ideology scores. Coppock hypothesized
that ideologically similar legislators would be more likely to share
information. If the constituent survey information does induce a change in behavior, this
effect may be observed not only in those legislators that are directly treated
in the experiment, but also by those of similar ideology to those that are
treated.

@@

Where $g$ is a function that normalizes the sums of distances to treated
neighbors to have unit variance. This model has a direct
effect ($\beta_1$) and an indirect effect ($\beta_2$) that is linear in the distances to
treated neighbors. Coppock evaluated this spillover model using the SSR+Design statistic as
presented in this paper. To demonstrate the value of this statistic,
particularly for models that are linear in the spillover effects, we repeat
Coppock's analysis using both the SSR+Design statistic and a KS statistic. The
results in Figure~\ref{fig:coppock-replication} recreate Coppock's Figure
2.\footnote{During the replication process, we noticed an inconsistency between
  our results using the SSR+Design statistic and Coppock's published figure. In
  coordination with the author, we discovered that the published figure suffered
  from on a programming error. The figure in this document reflects a correction
  to that error, so it will not precisely match Coppock's original publication.}
The KS test statistic, while being more discerning in the neighborhood of (0,
0), fails to bound the region of plausible hypotheses in the manner of the
SSR+Design statistic. We find this quality of the test statistic to be
particularly valuable. A powerful test statistic is a useful
tool, but equally useful is the ability to succinctly describe the region in
which we would fail to reject a null hypothesis. In this case, the SSR statistic
provides such a region, making discussing and evaluating the result's
of Coppock's analysis much more straightforward: it would be very unlikely that
the data were generated from the model with parameters outside the rectangle
with corners (-0.6, 0.2) and (0.3, -0.4).

\begin{figure}[h!] \centering
  \includegraphics[width=.99\textwidth]{../coppock-replication/CoppockJEPS_figure2.pdf}
  \caption{Replication of Figure 2 from \citet{coppock2014information}: plotting
    $p$-values for direct and indirect model parameters. The
    original analysis used the SSR statistic, which nicely bounds the region of
    plausible hypotheses. The KS statistic, while eliminating more hypotheses in
  the neighborhood of (0, 0), fails to provide any such bounds.}
  \label{fig:coppock-replication}
\end{figure}

@@

We cannot say here whether the SSR+Design test will provide the best power
against relevant alternatives for all possible models of treatment effect
propagation, network topologies and designs. However, we hope that this
research note both improves the application of the BFP approach and raises new
questions for research.  BFP are correct in the assertion that, regardless of
the choice of test statistic selection, a set of implausible hypotheses is
identified by the procedure. But we should not be led to believe that, for any
given test statistic, that some hypotheses are universally more plausible than others.
Such inferences --- comparing hypotheses --- may depend on the test statistic
used, and not necessarily reflect the plausibility of the model at hand. That
is, the results of any hypothesis test (or confidence interval creation) tell
us \emph{both} about the test statistic \emph{and} about the causal model under scrutiny.

In the example above, the SSR+Design test statistic had much better power than
any other test statistic. But SSR from an ordinary least squares regression is
not always appropriate: for example, when the probability of exposure to
spillover is heterogeneous across individuals in a way not well captured by the
$\bz^T \bS$ term or some other analogous term, we may wish to apply inverse
probability weights so as to ensure representative samples of potential
outcomes. This suggests a conjecture: that the $SSR$ from an \emph{inverse-probability-weighted} least squares regression is more generally a
sensible test statistic for models that include
interference.\footnote{\citet{aronowsamii2012interfere} use such weights for
  unbiased estimation of network-treatment-exposure probability weighted
  average treatment effects.}  Additionally, when nonlinear deviations from
model predictions are of concern, a weighted variant of the Brownian distance
covariance \cite{szekely2009brownian} or other {\it E}-statistic may be more
sensible than SSR.

>>>>

Notice that the test statistic choice matters in this process: the engine of
statistical inference involves summarizing information against the
hypothesized claim, yet different ways to summarize information might be more
or less sensitive to substantively meaningful differences. The statistical
power of a simple test of the sharp null hypothesis of no effects will vary as
a function of the design of the study (total number of observations, proportion treated, blocking structure,
etc), characteristics of the outcome (continuous, binary, skewed, extreme
points, etc), and the way that a test statistic summarizes the outcome (does
it compare means, standard deviations, medians, qqplots, etc). In general,
test statistics should be powerful against relevant alternatives. \citet[\S 2.4.4]{rosenbaum:2002} provides more specific
advice about the large sample performance of certain classes of test statistics
and BFP repeat his general advice: ``Select a test statistic [$\mathcal{T}$] that will be
small when the treated and control distributions in the adjusted data \ldots
are similar, and large when the distributions diverge.'' \cite[Proposition 4
and 5, \S 2.9]{rosenbaum:2002} presents results proving that test statistics
with this property (``effect increasing'' test statistics), produce an
unbiased test of the hypothesis of no effects or positive effects when the
positive effects involve one parameter. Notice that this means that, a test of
the sharp null of no effects
based on a known randomization using \emph{any} effect increasing
test-statistic to be valid test (in that the test should produce $p$-values
less than $\alpha$ no more than 100$\alpha$ of the time when the null is true) even though different
test statistics may imply different power against false hypotheses.  Yet, when the models are complex and
may involve increasing effects in the direction of one parameter and
non-linear effects in the direction of another parameter, BFP showed that
sometimes a KS-test will have (1) no power to address the model at all such
that all hypothesized parameters would receive the same implausibility
assessment; (2) or might reject all such hypothesized parameters. Thus,
although in theory one may assess sharp multiparameter hypotheses,
 in practice one may not learn much from such tests. BFP thus recommended simulation studies of the operating
characteristics of tests as a piece of their workflow --- because  the theory
justifying simple one-dimensional effect increasing test statistics clearly
did not cover multi-parameter situations like those easily arising from social
network experiments.

@@

Yet, one can also think about the process of hypothesis testing as a process
of assessing model fit, and there are usually better ways to evaluate the fit
of a model than comparing two marginal distributions. In the case where we know
the fixed adjacency matrix of the network, $\bS$, and where we imagine that
network attributes (like degree) of a node play a role in the mechanism by
which treatment propagates, the idea of assessing model fit rather than
closeness of distributions leads naturally to the sum-of-squared-residuals
(SSR) from a least squares regression of $ \widetilde y_{i,Z_i=0}$ on $Z_{i}$ and
$\bz^{T} \bS$ (i.e.\ the number of directly connected nodes assigned treatment)
as well as the $\mathbf{1}^{T} \bS$ (i.e.\ the degree of the node). If we
collected $Z_{i}$,$\bz^{T} \bS$, and $\mathbf{1}^{T} \bS$ into a matrix $\bX$,
and fit the $\widetilde y_{i,Z_i=0}$ as a linear function of $\bX$ with
coefficients $\bbeta$ when we could define the test statistic as:

\begin{equation}
 \T(\yu,\bz)_{\text{SSR}} \equiv \sum_{i} ( \widetilde y_{i,Z_i=0} - \bX\hat{\bbeta} )^2 \label{eq:ssr}
\end{equation}

To provide a little intuition about the SSR test statistic as compared to the
KS test statistic,

we created a small simulation study of 256 units with half
assigned to a binary treatment completely at random. We compared two outcomes
--- a Normal outcome and a skewed outcome, both with mean of 10 and standard
deviation of 1. And we compared two models of effects --- a constant additive
effects model in which $y_{i,Z_i=1}=y_{i,Z_i=0}+\tau$ and a constant
multiplicative effects model in which $y_{i,Z_i=1}=y_{i,Z_i=0}*\tau$. We set
the truth to be the sharp null of no effects $\tau=0$ for the additive model
and $\tau=1$ for the multiplicative model. Our intuition was that the SSR
would have more power than the KS test when the outcome was Normal and when
the observational implication of the model of effects would shift means (i.e.
the additive model). We used direct simulation of the randomization
distribution to generate $p$-values and repeated that process 1000 times to
gauge the proportion of rejections of false hypotheses (i.e. the power of the
tests). The results, not shown here but part of the reproduction archive for
this paper, bear out this intuition: the SSR has slightly more power than KS
for Normal outcomes in both the additive and multiplicative effects
conditions. SSR has slightly less power than KS when the outcome is skewed for
both models. In general, the SSR ought to be most powerful when the effect of
the experiment involves a shift in the location of the distributions of the
treated versus the controls. And it should be less powerful when the effect of
the experiment is mostly to leave the center of the distribution alone, but,
instead to stretch or compress the tails, or to concentrate the experimental
effect on one quantile.

\subsection{The SSR Test Statistic and the BFP Example Model}

@@

\citet{coppock2014information} presents an reanalysis of an experiment performed
by \citet{butler2011can}. Leading up to a key vote on a budget bill, SB24, in the
New Mexico state senate, Butler and Nickerson randomly assigned legislators to see
constituent survey results. The original
analysis found that legislators assigned to receive constituent information were
more likely to vote consistently with the preferences of their district.
Given the easy nature of sharing information in
field experiments \citep{winters2012using},  Coppock reanalyzes the
experimental results with an eye towards spillover, constructing a social
network based on similarity on W-NOMINATE ideology scores. Coppock hypothesized
that ideologically similar legislators would be more likely to share
information. If the constituent survey information does induce a change in behavior, this
effect may be observed not only in those legislators that are directly treated
in the experiment, but also by those of similar ideology to those that are
treated.

@@

Where $g$ is a function that normalizes the sums of distances to treated
neighbors to have unit variance. This model has a direct
effect ($\beta_1$) and an indirect effect ($\beta_2$) that is linear in the distances to
treated neighbors. Coppock evaluated this spillover model using the SSR+Design statistic as
presented in this paper.

\begin{figure}[H] \centering
  \includegraphics[width=.75\textwidth]{../coppock-replication/CoppockJEPS_figure2.pdf}
  \caption{Replication of Figure 2 from \citet{coppock2014information}: plotting
    $p$-values for direct and indirect model parameters. The
    original analysis used the SSR statistic, which nicely bounds the region of
    plausible hypotheses. The KS statistic, while eliminating more hypotheses in
  the neighborhood of (0, 0), fails to provide any such bounds.}\label{fig:coppock-replication}
\end{figure}

To demonstrate the performance of this statistic in
an applied setting, such as this one in which  the model is linear in the spillover effects, we repeat
Coppock's analysis using both the SSR+Design statistic and the simple KS statistic. The
results in Figure~\ref{fig:coppock-replication} recreate Coppock's Figure
2.\footnote{We are grateful to Alex Coppock for his help with both code and
  data. Figure~\ref{fig:coppock-replication} does not precisely match the
  figure in the original publication because we corrected a small
  programming error during replication.}
 The KS test statistic, while being more powerful in the neighborhood of (0,
0), fails to bound the region of plausible hypotheses in the manner of the
SSR+Design statistic. We find this quality of the test statistic to be
particularly valuable. A powerful test statistic is a useful
tool, but equally useful is the ability to succinctly describe the region in
which we would fail to reject a null hypothesis. In this case, the SSR statistic
provides such a region, making discussing and evaluating the result's
of Coppock's analysis much more straightforward: it would be very unlikely that
the data were generated from the model with parameters outside the rectangle
with corners (-0.6, 0.2) and (0.3, -0.4).

@@

We cannot say here whether the SSR+Design test will provide the best power
against relevant alternatives for all possible models of treatment effect
propagation, network topologies and designs: this paper uses two models of
effects each of which was applied to a different network topology and
experimental design.\footnote{The BFP paper itself engages with some questions about
the performance of this approach when the theoretical model is very different
from the process generating the data, and we encourage readers to see that
discussion in their \S~5.2.} However, we hope that the results from the 
examples presented here improves the application of the BFP approach and raises new
questions for research.  BFP are correct in the assertion that, regardless of
the choice of test statistic selection, a set of implausible hypotheses is
identified by the procedure. But we should not be led to believe that, for any
given test statistic, that some hypotheses are universally more plausible than others.
Such inferences --- comparing hypotheses --- may depend on the test statistic
used, and not necessarily reflect the plausibility of the model at hand. That
is, the results of any hypothesis test (or confidence interval creation) tell
us \emph{both} about the test statistic \emph{and} about the causal model under scrutiny.

In the example shown in Figure~\ref{fig:twoD}, the SSR+Design test statistic had much better power than
any other test statistic. But SSR from an ordinary least squares regression is
not always appropriate: for example, when the probability of exposure to
spillover is heterogeneous across individuals in a way not well captured by the
$\bz^T \bS$ term or some other analogous term, we may wish to apply inverse
probability weights so as to ensure representative samples of potential
outcomes. This suggests a conjecture: that the $SSR$ from an \emph{inverse-probability-weighted} least squares regression is more generally a
sensible test statistic for models that include
interference.\footnote{\citet{aronowsamii2012interfere} use such weights for
  unbiased estimation of network-treatment-exposure probability weighted
  average treatment effects.}  Additionally, when nonlinear deviations from
model predictions are of concern, a weighted variant of the Brownian distance
covariance \citep{szekely2009brownian} or other $E$-statistic may be more
sensible than the sum of squared residuals.

### 2015-12-19 paper/paper.tex
More progress on editing

%\titlespacing*{ ⟨command⟩}{⟨left⟩}{⟨beforesep⟩}{⟨aftersep⟩}[⟨right⟩]
\titlespacing{\section}{0pc}{1.5ex plus .1ex minus .2ex}{.5ex plus .1ex minus .1ex}
\titlespacing{\subsection}{0pc}{1.5ex plus .1ex minus .2ex}{.5ex plus .1ex minus .1ex}
\titlespacing{\subsubsection}{0pc}{1.5ex plus .1ex minus .2ex}{.5ex plus .1ex minus .1ex}

\newif\ifanon

% \ifanon
% \footnote{We are awesome!}
% \else
% \fi
%% \abovedisplayshortskip=0pt
%% \belowdisplayshortskip=0pt
%% \abovedisplayskip=0pt
%% \belowdisplayskip=0pt
%% 
\begin{document}
%\VerbatimFootnotes

>>>>

%\titlespacing*{ ⟨command⟩}{⟨left⟩}{⟨beforesep⟩}{⟨aftersep⟩}[⟨right⟩]
\titlespacing{\section}{0pc}{1.5ex plus 0.1ex minus 0.2ex}{.5ex plus 0.1ex minus 0.1ex}
\titlespacing{\subsection}{0pc}{1.5ex plus 0.1ex minus 0.2ex}{.5ex plus 0.1ex minus 0.1ex}
\titlespacing{\subsubsection}{0pc}{1.5ex plus 0.1ex minus 0.2ex}{.5ex plus 0.1ex minus 0.1ex}

%%% Links and Table of Contents
\usepackage[dvipsnames]{xcolor}
\usepackage[colorlinks=TRUE,linkcolor=BrickRed,citecolor=MidnightBlue]{hyperref}

% \newif\ifanon
% \ifanon
% \footnote{We are awesome!}
% \else
% \fi
%% \abovedisplayshortskip=0pt
%% \belowdisplayshortskip=0pt
%% \abovedisplayskip=0pt
%% \belowdisplayskip=0pt
%% 
\begin{document}
%\VerbatimFootnotes
